---
title: "species_assemblage_networks"
author: "Otis Brunner"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 5
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '5'
---


```{r setup, include=FALSE}
# This saves all figures in the directory `figures` and sets the default figure size 
knitr::opts_chunk$set(echo = TRUE, fig.path = 'figures/')
```



This project is linked to the following Github repository: https://github.com/otistwo/hydrothermal_thesis.git


## Load Libraries

```{r libraries, results="hide"}
library(data.table)
library(vegan)
library(ggplot2)
library(clustsig)
library(reshape2)
library(igraph)
library(plot.matrix)
library(sidier)
library(pals)
library(qdapTools)
library(marmap)
library(readr)
library(tidyverse)
library(rgdal)
library(raster)
library(rnetcarto)
library(sp)


# Load the required packages
# (vegan must be loaded after ade4 to avoid some conflicts)
library(ape)

#install.packages("packfor", repos="http://R-Forge.R-project.org")
#library(packfor)

library(tripack)
#install.packages("spacemakeR", repos="http://R-Forge.R-project.org")
library(spacemakeR)

library(ade4)
library(adespatial)
library(spdep)
library(vegan)

#install.packages("AEM", repos="http://R-Forge.R-project.org")
library(AEM)

#install.packages("PCNM", repos="http://R-Forge.R-project.org")
#library(PCNM)

library(SoDA)

# Load additionnal functions
# (files must be in the working directory)
source("plot.links.R")
source("sr.value.R")
source("quickPCNM.R")

```
# The Data

## Importing and formatting 

Import the presence/absence dataset fromed from BISMaL + Nakajima et al. (2014) + Watanabe et al. (2019) + Tunnicliffe (unpublished) + Giguere (unpublished) 
```{r import site by species presence}
pres <- read.csv("data/site_by_species_presence.csv")

#order data-set by vent names for simplicity
pres <- with(pres,  pres[order(vent) , ])
```

remove the completely disconnected 'kueshan' and the shallow water vents 'Omura' and 'Kagoshima'
```{r}
pres <- pres[pres$vent!="Kueishan", ]
pres <- pres[pres$vent!="Omuro", ]
pres <- pres[pres$vent!="Kagoshima", ]

pres.meta<- pres[,120:139]
rownames(pres.meta)<-pres$vent
```

```{r subset the pres/abs matrix}
mat <- as.matrix(pres[,3:119])
rownames(mat) <- pres$vent
```

# Similarity Analyses

Calculate Sorensen's Coefficient to determine pairwise site dissimilarity. Sorensen's Coefficient gives double weight to shared presence over shared absence

```{r Sorensens coefficient Similarity Profile}
#Calculate Sorensens Coeeficient 
vent.dist <- vegdist(mat, method = "bray", diag = TRUE, alpha = 0.05)
```

## SIMPROF Clusters
```{r SIMPROF}
#Cluster communities using SIMPROF and Bray-curtis
simpr <- simprof(mat, method.distance = "czekanowski", num.expected = 100, num.simulated = 999, alpha = 0.05)

#extract the cluster membership
clusts<-melt(simpr$significantclusters)

#rename column names
colnames(clusts)<- c("Name","clusts")

#Attach cluster identity to the pres data-set
pres<-merge(pres, clusts, by.x = "vent", by.y = "Name")
```

Standardise a colour pallete based on the individual clusters for use in later plots for consistency 
```{r SIMPROF Clustering with Sorensen}
#Create a colour pallette
cp.clusts=c( glasbey(length(unique(clusts$clusts))))[clusts$clusts]

p1r <- simprof.plot(simpr, plot = TRUE, leafcolors = unique(cp.clusts))
```

Check how depth and position relate to the clustering

```{r cluster by IR parameters}

#Plot the vents coloured by clustering
plot( pres$Latitude, pres$Maximum.or.Single.Reported.Depth, main = "Simprof clustering", xlab = "Lat", ylab = "Depth", col = cp.clusts, pch=16, cex=2.5 )
text(pres$Latitude, pres$Maximum.or.Single.Reported.Depth,  labels=pres$vent,  pos=3, offset=0.5)

#ggplot (param, aes(x = Lattitude, y = Depth, color =as.factor (clusts))) + geom_point() + scale_colour_brewer("glasbey")
  
  
```

# Similarity Network Formation

reformat the dissimilarity matrix to work with igraph package

```{r Reformating the dissimilarity matrix}
#use Similarity to represent strength of connections
vent.sim <- 1 - vent.dist

net<-graph_from_adjacency_matrix(as.matrix(vent.sim), mode = "undirected", weighted = TRUE)
```

Plot the network with default plot settings

```{r  unedited similarity network}
#Plot NW. Pacific Metacommunity
plot(net)
```

## Editing graph attributes

At first the network has no attributes other than the vertex names
```{r }
vertex_attr_names(net)
```
 add attributes of significance from the InterRidge data 
```{r}
#Add from param (excluding name)
V(net)$lat<-pres$Latitude
V(net)$long<-pres$Longitude
V(net)$depth<-pres$Maximum.or.Single.Reported.Depth
V(net)$tect<-pres$Tectonic.setting
V(net)$region<-pres$Region
V(net)$clusts<-pres$clusts 

vertex_attr_names(net)

attri_table<- as.data.frame(vertex.attributes(net))
```

Plot the network but colour the vertexes by their SIMPROF clustering
```{r Similarity Network coloured by SIMPROF cluster}

cp.clusts=c( glasbey(length(unique(V(net)$clusts))))[(V(net)$clusts)]

#Use colour pallete for clustering
plot(net, edge.curved=.5, margin = 0, edge.arrow.size=.1, abel.dist=1.5, vertex.color=cp.clusts)

#Create legend for region
legend(x=1.3, y=1., unique(V(net)$clusts), col =unique(cp.clusts), pch=19, cex = 1, bty="n")
```
Plot the network but colour the vertexes by their defined region
```{r Similarity Network coloured by region }

#Create a colour pallette for region
cp.region=c( glasbey(length(unique(V(net)$region))))[(as.factor(V(net)$region))]

plot(net, edge.curved=.5, margin = 0, edge.arrow.size=.1, abel.dist=1.5, vertex.color=cp.region)

#Create legend for region
legend(x=1.3, y=1., unique(V(net)$region), col =unique(cp.region), pch=19, cex = 1, bty="n")
```

## Thresholded dissimilarity network

Calculate the Percolation Threshold of the similarity network censu Rozenfeld et al. (2008)
```{r, Percolation threshold under single NW. Pacific metacommunity hypothesis, results="hide"}

#Transform Sorensen's coefficient matrix into a data table
vent.dist <- as.matrix(vent.dist)
vent.dist <- as.data.frame(vent.dist)

percolations <- perc.thr(vent.dist, range = seq(0, 1, 0.001))
```

```{r Percolation Threshold identification}
perc.sum<-percolations$Summary
perc.sum<-as.data.frame(perc.sum)

plot(perc.sum[,1], perc.sum[,2], )

ggplot(perc.sum, aes( x = Threshold, y = `<s>`)) +
  geom_point(size = 0.5) +
  labs(x = "Threshold", y = "S") +
  ggtitle("Percolation Threshold") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) 
  
```

```{r  Similarity network with threshold applied}

#Use the calculated percolation threshold to cut the edges of the graph. As the threshold is calculated using dissimilarity and igraph uses similarity the cutting value is the inverse of the percolation threshold. (AKA Keep all edges that represent a similarity lower than the maximum dissimilarity threshold)
perc.net <- igraph::delete.edges(net, which(E(net)$weight < 1 - percolations$`Estimated Percolation Threshold`))

plot(perc.net, edge.curved=.5, ,vertex.label.color="gray28", vertex.size = 8, vertex.label.cex=0.7, vertex.label.font = 2, vertex.label.dist = 0.8, vertex.label.degree = -pi/2,  edge.arrow.size=.1, abel.dist=1.5, vertex.color=cp.clusts, layout = layout_with_kk)

```

```{r Thresholded Similarity network by tectonic setting}

plot(perc.net, edge.curved=.5, ,vertex.label.color="gray28", vertex.size = 8, vertex.label.cex=0.7, vertex.label.font = 2, vertex.label.dist = 0.8, vertex.label.degree = -pi/2,  edge.arrow.size=.1, abel.dist=1.5, vertex.color=cp.region, layout = layout_with_kk)
```

# Similarity Network Statistics

## Betweenness Centrality

```{r Betweenness Centrality - Thresholded similarity network by region and sized }
#Calculate Betweeness Centrality for thresholded networ
b.p.cent<-igraph::betweenness(perc.net, directed = FALSE, weights = NA)

plot(perc.net, edge.curved=.5, ,vertex.label.color="gray28", vertex.size = b.p.cent/8, vertex.label.cex=0.7, vertex.label.font = 2, vertex.label.dist = 0.8, vertex.label.degree = -pi/2,  edge.arrow.size=.1, abel.dist=1.5, vertex.color=cp.region, layout = layout_with_kk)
```

# Bipartite Network Formation

Trasform the presence/absence matrix into a Bipartite network were vertices that represent species are linked to vertices that represent vent sites by an unweighted, un-directed edge.

```{r reformat matrix}
blinks <- melt(as.matrix(mat), varnames  = c("Source", "Target"))

# remove all rows with weight of "0" (no presence)
blinks <- blinks[apply(blinks!=0, 1, all),]

#Remove the value column so that you simply have a list of connected nodes
blinks <- blinks[1:2]

```

```{r simple bipartite network, results="hide"}

#Transform it into a graph object
bgraph <- graph.data.frame(blinks, directed=FALSE)

#Evaluate graph object to see if it meets the criteria of a 2-node network and the 'type' used to identify between node types (sites/species)
bipartite.mapping(bgraph)

V(bgraph)$type <- bipartite_mapping(bgraph)$type

V(bgraph)$color <- ifelse(V(bgraph)$type, "lightblue", "salmon")

plot(bgraph)

```

```{r, Bipartite layout }

plot(bgraph, layout=layout.bipartite,vertex.size=3, vertex.label=NA)
```

# Bipartite Network Statistics

## Modularity

Run netcarto on the adjacency matrix of vents and species
```{r bipartite node details for gephi}
#return the list of nodes and their type (species/site)
bnodes <- igraph::as_data_frame(bgraph, what = "vertices")
#blinks <- merge(blinks, bnodes, by.x = "Source", by.y = "name")

#rename the type to vents or species
bnodes$type <- as.character(bnodes$type)
bnodes$type[bnodes$type == "FALSE"] <- "vent"
bnodes$type[bnodes$type == "TRUE"] <- "species"

bmods<-netcarto(blinks, seed = as.integer(floor(runif(1, 1, 10000001))), iterfac = 100,
  coolingfac = 0.995, bipartite = FALSE)

```

Append the netcarto scores to the bgraph graph object and plot with module membership as colours
```{r bipartite network with modules}
#Convert the netcarto results to a data frame
bmods<-as.data.frame(bmods)

#extract the nodelist as it is in the graph object into a column
bgraph.nodes<-as.list(V(bgraph))
bgraph.nodes<-t(as.data.frame(bgraph.nodes))
bgraph.nodes<- rownames(bgraph.nodes)
bgraph.nodes<- as.data.frame(bgraph.nodes)

#Combine module scores with the nodelist so that they are in the same order (sort = FALSE)
bgraph.modsc <- merge(bgraph.nodes, bmods, by.x = "bgraph.nodes", by.y = "name",  sort = FALSE )

#Add a 1 to module scores as some are given 0
bgraph.modsc$module<-bgraph.modsc$module+1

#Insert the module scores as a vertex attribute in the graph
V(bgraph)$mods <- bgraph.modsc$module

#Match the colours from gephi
g.col <- c("orange","lightblue","green","purple")
V(bgraph)$color <- g.col[V(bgraph)$mods]

plot(bgraph, vertex.size = 10, vertex.label.cex = 0.5, layout = layout_with_kk)

```

## Within/Between Module Degree 

```{r vent site modularity degree}

#Trim to only include vent nodes
vent.modsc<- bgraph.modsc[1:32,]

#Match the colours from gephi
g.col <- c("purple","lightblue","green","orange")
v.cols <- g.col[vent.modsc$module]

plot(vent.modsc$participation, vent.modsc$connectivity, main = "NETCARTO Vent Node Importance", xlab = "Participation Coefficient", ylab = "Within Module Degree", , col=v.cols, xlim = c(0,1), ylim = c(-0.5,5), pch = 16)

text(vent.modsc$participation, vent.modsc$connectivity, labels = vent.modsc$bgraph.nodes, cex=0.7, pos=3, offset=0.2)


abline(h = 2.5)
abline(v= 0.62)

```

This shows the role of each vent site in terms of being a hub of connectivity for their module (Within Module Degree) and their role in connecting between modules (Participation Coeffiecient). The lines delineate four distinct roles based on these values (Guimera and Amaral, 2005)

```{r, Species modularity degree}

#Trim to only include vent nodes
species.modsc<- bgraph.modsc[32:146,]

#Match the colours from gephi
g.col <- c("orange","lightblue","green","purple")
v.cols <- g.col[species.modsc$module]

plot(species.modsc$participation, species.modsc$connectivity, main = "NETCARTO Species Node Importance", xlab = "Participation Coefficient", ylab = "Within Module Degree", xlim = c(-0.1,1), ylim = c(-1,5),pch=16)

text(species.modsc$participation, species.modsc$connectivity, labels = species.modsc$bgraph.nodes, cex=0.7, pos=3, offset=0.2)


abline(h = 2.5)
abline(v= 0.62)

```

# Export Network Outputs

Export the characteristics of each vent site node in terms of the statistics carried out on the similarity- and bipartite networks.

```{r}
vent.nodes<-merge(attri_table, bmods, by.x = "name", by.y = "name")
vent.nodes<-vent.nodes[-12]

str(vent.nodes)
```
```{r}
write.csv(vent.nodes, "vent_nodes.csv", row.names = FALSE)

```

## Export for use in 'Gephi' GUI

```{r}

```


# Post-hoc analysis

Subset the datasets for analyses following the methods outlined in 'Numerical Ecology with R'

## Formatting
```{r}

vent<- as.data.frame(mat)
vent.env<- vent.nodes[c(4:6)]
vent.ll<- vent.nodes[c(3,2 )]

#Translate degree coordinates into Cartesian with the resultant unit being km
vent.xy<- geoXY(vent.ll$long,vent.ll$lat, unit = 1000)
vent.xy<-as.data.frame(vent.xy)
vent.xy<-vent.xy[c(2,1 )]

# Transform the data
vent.h <- decostand (vent, "hellinger")
vent.xy.c <- scale(vent.xy, center=TRUE, scale=FALSE)

```


## Depth correlogram


Spatial correlogram (based on Moran's I)

 Search for neighbours of all points within a radius of 2100km and multiples (i.e., 0 to 300km, 300 to 600 km, ...). Sites form a fully connected network at ~1360kmand above (below which Okinawa Trough sites are seperate).
```{r univariate correlogram}

quartz(title="Linkage map")

#12.7 = percolation threshold
nb1 <- dnearneigh(as.matrix(vent.xy), 0, 300)
summary(nb1)

```

```{r linkage at 300km}
plot.links(vent.xy, thresh=300)

#1361
```

```{r Correlogram of depth}

depth <- vent.env$depth
depth.correlog <- sp.correlogram(nb1, depth, order=7, method="I", style = "W",	zero.policy=TRUE)
print(depth.correlog, p.adj.method="holm")
quartz(title="Correlogram of Depth")

```
```{r  depth correlog}
plot(depth.correlog)
```
Depth Correlogram Interpretation: There is no significant correlation between depth and distance at any of the scales so can be considered spatially independent in regards to depth.

## Multivariate Correlogram of species composition

Run with default settings (distance class determined by Sturges equation)
```{r  community correlogram}
(vent.correlog <- mantel.correlog(vent.dist, XY=vent.xy, nperm=999))
summary(vent.correlog)
```
Check the distance classes being used
```{r community correlog features}
# Number of classes
vent.correlog$n.class # or: vent.correlog[2]
# Break points
vent.correlog$break.pts # or: vent.correlog[3]
```
Multiply distance classes by 100 to get the value in kilometers
```{r Community correlogram}

#quartz(title="Mantel correlogram of vent data")
plot(vent.correlog)
```
There is a significant positive correlation between site dissimilarity and distance at and below distance class 2 (when vents at and below 263 km distance apart are connected). Below is what this network would look like:

```{r links at significant distance class}
plot.links(vent.xy, thresh=263)
```
At a much greater distance of 1045km the difference between sites become negatively correlated to distance. This means that sites a number of this size step away are more similar than those just one step away at this distance
```{r links at non-significant distance class}
plot.links(vent.xy, thresh=1045.155279)
```

Run again but with custom distance classes
```{r}
bp<- seq(300, 2100, length.out=7)
```

```{r, warning=FALSE, message=FALSE}
(vent.correlog <- mantel.correlog(vent.dist, XY=vent.xy, nperm=999, break.pts = bp))
summary(vent.correlog)
```
Check the distance classes being used
```{r mantel custom features}
# Number of classes
vent.correlog$n.class # or: vent.correlog[2]
# Break points
vent.correlog$break.pts # or: vent.correlog[3]
```

```{r multivariate correlogram with custom distance classes}

#quartz(title="Mantel correlogram of vent data")
plot(vent.correlog)
```
The results are very similar

## Distance-based Moran's Eigenvector Maps

dbMEMs are a way to distill the geographical position of each site into a parameter that can be used in further analysis that aren't able to compare matrices. 

First re-order the data-sets so that they are in two groups, Oki vents first and then the rest. This is necessary to create a dbMEM that does not connect across the two groups (informed by above analyses)
```{r re-order data-sets}

#re order vents so that the first 8 are OKI vents (re order by longitude)
basin.pres  <- with(pres,  pres[order(Longitude) , ])
basin.mat <- as.matrix(basin.pres[,3:119])
basin.vent<- as.data.frame(basin.mat)

basin.vent.nodes<- with(vent.nodes,  vent.nodes[order(long) , ])

basin.vent.env<- basin.vent.nodes[c(4:6)]
basin.vent.ll<- basin.vent.nodes[c(2,3 )]


```

```{r create the dbMEM}
#Dictate the number of sites in each spatial group (Oki and I-B-M)
nsites.per.group <- c(8,24)

dbmem2<-create.dbMEM.model(coord = basin.vent.ll, D.mat = NULL, nsites = nsites.per.group)

dbmem1<-create.dbMEM.model(coord = basin.vent.ll, D.mat = NULL, nsites = 32)

dbmem2


```
Plot the groups as a map to make sure it's correct

```{r dbMEM groups}
# Draw a map to check the coding of the sites into the groups
 site.codes <- unlist(apply(cbind(1:2),1,n=nsites.per.group,function(a,n) rep(a,n[a])))

 ggplot(basin.vent.ll, aes(x = long, y = lat)) +
   geom_point(aes(color = site.codes))
```
## Redundancy Analysis

```{r redundancy analysis}

vent.rda<- rda(basin.vent,   dbmem2)
vent.rda
(R2 <- RsquareAdj(vent.rda)$r.squared)
```


## Variance Partitioning

```{r variance partitioning betwween depth(X1), fine-spatial(X2), broad-spatial (X3)}
# Partition Bray-Curtis dissimilarities
varp_vent<-varpart(vegdist(basin.vent), basin.vent.env$depth, dbmem2, dbmem1)

varp_vent
plot(varp_vent)
plot (varp_vent, digits = 1, Xnames = c('depth', 'f-spatial', 'b-spatial'), bg = c('blue', 'tomato', 'green'))
```

```{r variance partitioning between depth and tectonic setting(X1), fine-spatial(X2), broad-spatial(X3)}
# Partition Bray-Curtis dissimilarities
varp_vent<-varpart(vegdist(basin.vent), basin.vent.env[1:2], dbmem2, dbmem1)

varp_vent
plot (varp_vent, digits = 1, Xnames = c('depth & tectonic', 'f-spatial', 'b-spatial'), bg = c('blue', 'tomato', 'green'))
```


```{r variance partitioning between depth (X1) and fine-spatial(X2)}
# Partition Bray-Curtis dissimilarities
varp_vent<-varpart(vegdist(basin.vent), basin.vent.env$depth, dbmem2)

varp_vent
plot (varp_vent, digits = 1, Xnames = c('depth ', 'f-spatial'), bg = c('blue', 'tomato'))
```

```{r variance partitioning between depth (X1) and broad-spatial(X2)}
# Partition Bray-Curtis dissimilarities
varp_vent<-varpart(vegdist(basin.vent), basin.vent.env$depth, dbmem1)

varp_vent
plot(varp_vent, main = "variance partitioning between depth (X1) and broad-spatial(X2)")
```