---
title: "data_creation"
author: "Otis Brunner"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 5
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '5'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = 'figures/')
```

```{r libraries, results='hide'}
library(sp)
library(rgdal)
library(dplyr)
library(maptools)
library(maps)
library(devtools)
library(ggmap)
library(marmap)
library(sf)
library(raster)
library(readr)
library(reshape)
library(spaa)
library(fuzzySim)
library(readxl)
library(plyr)
library(worms)
```
# Data-sets

**Data-sets all downloaded on 2020/11/26**

BISMaL data downloaded at https://www.godac.jamstec.go.jp/bismal/e/occurrences
See README for details on how it was downloaded (Latittude: 6 - 36; Longitude: 120 - 150)

InterRidge dataset downloaded from https://doi.pangaea.de/10.1594/PANGAEA.917894?format=html#download

SFDVENT database downloaded at https://onlinelibrary.wiley.com/doi/full/10.1111/geb.12975

```{r Import}

bis_origi <- read.csv("data/BISMaL_DwC20201126101643.csv")

taxa_origi <- read.csv("data/geb12975-sup-0007-supinfo2.csv")

vent_origi <- read.csv("data/vent_fields_all_20200325cleansorted.csv")
```

##Clean columns

Remove variables (columns) from original data-sets that aren't needed
```{r Trim Columns}
#Create new datasets with only target variables (columns)
taxa <- taxa_origi[c ("Taxon","Family", "Genus","Species")]


bis <- bis_origi [c ("scientificName","family","genus","specificEpithet", "locality", "decimalLatitude", "decimalLongitude", "eventDate", "fieldNumber", "institutionCode", "datasetName", "basisOfRecord", "references", "occurrenceID", "catalogNumber", "recordNumber", "recordedBy", "BISMaLtaxonID")]

vent <- vent_origi [c ("Name.ID", "Latitude" , "Longitude", "Region", "Maximum.or.Single.Reported.Depth" , "Tectonic.setting")]

```
# Spatialize

## Map BISMaL samples

Transform the bismal dataframe into spatial points that can be plotted against a basemap of NOAA Bathymetry
```{r Spatialise BISMaL data}
coords_bis<- bis [c("decimalLongitude","decimalLatitude")]
sd_bis <- st_as_sf(bis, coords = c("decimalLongitude","decimalLatitude"), crs = 4326)
```

Create a basemap from NOAA bathymetry, designating the region of interest, resolution and colour pallete
```{r Basemap creation}
#First Create a Basemap using NOAA bathymetry data
bmap<-marmap::getNOAA.bathy(118, 150, 12, 35, resolution = 2)

# Creating a custom palette of blues
blues <- c("lightsteelblue4", "lightsteelblue3",
"lightsteelblue2", "lightsteelblue1")
```



Plot all of the sample points from the downloaded Bismal dataset against the basemap
```{r Map original BISMal points}

# Plotting the bathymetry with different colors for land and sea
plot(bmap, image = TRUE, land = TRUE, lwd = 0.1,
bpal = list(c(0, max(bmap), "darkgreen"),
c(min(bmap),0,blues)))

# Making the coastline more visible
plot(bmap, deep = 0, shallow = 0, step = 0,
lwd = 0.4, add = TRUE)

#Add the Bismal sample points to the basemap
points(coords_bis, cex = 0.2, pch = 16)
```
While Longitudinal transects are clear, it is also possible to see clustering of sample points along volcanic arcs and back-arcs where targeted hydrothermal vent sampling has been carried out


## Map InterRidge 3.4

Now add the information from InterRidge by also transforming that dataframe into spatial points
```{r Spatialise InterRidge data}
coords_IR<- vent [c("Longitude","Latitude")]
sd_IR <- st_as_sf(vent, coords = c("Longitude","Latitude"), crs = 4326)
```

The InterRidge data includes global vents so trim them to the extent of the NWP manually to exlude all vent points outside of the NWP
```{r match IR and bis}
e <- extent(raster(xmn = 118, xmx = 150, ymn = 10, ymx = 35))
sd_IR<-st_crop(sd_IR, e)
st_crs(sd_IR) = st_crs(sd_bis)
```

Now plot the vent sites from InterRidge and sample locations from Bismal together over the NOAA basemap
```{r Map of original Bis data and Vent Sites, NWP}

plot(bmap, image = TRUE, land = TRUE, lwd = 0.1,
bpal = list(c(0, max(bmap), "darkgreen"),
c(min(bmap),0,blues)))
# Making the coastline more visible
plot(bmap, deep = 0, shallow = 0, step = 0,
lwd = 0.4, add = TRUE)

plot(st_geometry(sd_IR), col = sf.colors(12, categorical = TRUE), border = 'grey', 
     axes = TRUE, add = TRUE)
plot(st_geometry(st_centroid(sd_IR)), pch = 3, col = 'red', add = TRUE)
plot(sd_bis,pch = 16, cex = 0.1, col = 'black', add=TRUE)
```

Now plot using ggplot to give more flexibility

```{r Map N.Oki vent sites labelled}
library(ggplot2)
library(mapdata)
#devtools::install_github("yutannihilation/ggsflabel")
library(ggsflabel)
library(gghighlight)


# convert bathymetry to data frame
bf = fortify.bathy(bmap)

# get regional polygons
reg = map_data("world2")
reg = subset(reg, region %in% c('Japan', 'Mariana', 'China'))


# set map limits 
lons = c(118, 150)
lats = c(10, 35)



# make plot
ggplot()+

     
    # add coastline
  geom_polygon(data = reg, aes(x = long, y = lat, group = group), 
               fill = "darkgreen", color = NA) + 
  
   # add 100m contour
  geom_contour(data = bf, 
               aes(x=x, y=y, z=z),
               breaks=seq(-100, -2000,- 150),
               size=c(0.3),
               colour="lightgrey")+
  
  
  

  # add vent points
  geom_point(data = vent, aes(x = Longitude, y = Latitude),
             colour = "black", fill = "red", 
             stroke = .5, size = 3, 
             alpha = 1, shape = 21)+
  
    # add sample points
  geom_point(data = bis, aes(x = decimalLongitude, y = decimalLatitude),
             colour = "orange", 
             stroke = .1, size = 1, 
             alpha = .1, shape = 16)+
  #Label the vents
  geom_label(data = vent, aes(x = Longitude, y = Latitude, label = Name.ID),
             size = 2,
             fontface = "bold",
             nudge_y = 0.1) +
  
  #add sample labels but be careful not to include too many samples by for example only labelling a subset of the data that is in the region "Okinawa Trough"
  geom_sf_label_repel(data = subset(sd_IR, Region  == "Okinawa Trough"), aes(label = Name.ID),
             size = 1,
             fontface = "bold",
             nudge_y = 0.1) +
  


  #focus on a region of interest e.g. Higashi-Ensei
    #coord_sf(xlim = c(126, 129), ylim = c(26, 29), expand = FALSE) +
     
  
   #focus on a region of interest e.g. Yokosuka
    coord_sf(xlim = c(125, 128.7), ylim = c(25.5, 29), expand = FALSE) +
      
    theme_classic() +
  
  labs(
    x = "Long.",              # x axis title
    y = "Lat."  # y axis title

  )+

 
  
  ggtitle("Northern Okinawa InterRidge Vents and Bismal samples")+
    theme(plot.title = element_text(hjust = 0.5))

```


## Create a Buffer around InterRidge vent sites

To match occurences from Bismal to vent sites we will do this by spatially joining the data from  InterRidge with any Bismal samples that occur within 5km of the InterRidge vent records.

To create a 5km buffer, we need to project the lat/long of the bismal and IR points into WGS84 using zone 52N (roughly central to the study region). Additionally the other spatial features of country shapes and bathymetry lines need to be projected
```{r Transform spatial points to metric coordinate system}

IR_km = st_transform(sd_IR, "+proj=utm +zone=52N +datum=WGS84 +units=km")
bis_km = st_transform(sd_bis, "+proj=utm +zone=52N +datum=WGS84 +units=km")

#Convert Bathymetry contours and country shapefiles for plotting

#Countries
sd_reg<-st_as_sf(reg, coords = c("long","lat"), crs = 4326)
#Project into WGS84
reg_km = st_transform(sd_reg, "+proj=utm +zone=52N +datum=WGS84 +units=km")
#Transform 'geometry' back into 'x' 'y' columns
reg_xy <- do.call(rbind, st_geometry(reg_km)) %>% 
     as_tibble() %>% setNames(c("x","y"))

reg_xy <-  cbind(reg_km[,3],reg_xy)

#Bathymetry
sd_bf<-st_as_sf(bf, coords = c("x","y"), crs = 4326)
#Project into WGS84
bf_km = st_transform(sd_bf, "+proj=utm +zone=52N +datum=WGS84 +units=km")
#Transform 'geometry' back into 'x' 'y' columns
bf_xy <- do.call(rbind, st_geometry(bf_km)) %>% 
     as_tibble() %>% setNames(c("x","y")) 

 bf_xy <-  cbind(bf_km[,1],bf_xy)

```

Transformations can often distort the data so it is important to check that they still look the same when plotted

```{r Map UTM transformation}
plot(st_geometry(IR_km), col = sf.colors(12, categorical = TRUE), border = 'grey', 
     axes = TRUE)
plot(st_geometry(st_centroid(IR_km)), pch = 3, col = 'red', add = TRUE)
plot(bis_km,pch = 16, cex = 0.1, col = 'black', add=TRUE)
```

The Bismal samples still outline the japanese coast and follow the Mariana-Izu-Bonin Arc and the InterRidge vent sites still occur along tectonic arcs and back arcs, overlapping with Bismal

Now create a circle buffer with a 5km buffer around each IR vent location. Then combine the data from the IR for the vent wth the occurence data from all Bismal samples located within the buffer

```{r Create vent buffer}

#Create a 5km Buffer
IR_buffer = st_buffer(IR_km, 5)

#Trim the BISMaL samples to only include those within a vent buffer (Just like `Spatial Join in Arc GIS)
IR_bis = st_intersection(IR_buffer, bis_km)

```
Check how many unique spaecies and vent sites are retained within this data-set that has removed all Bismal samples outside of the buffer
```{r summary of buffered datase}
summary(unique(IR_bis$scientificName))
summary(unique(IR_bis$Name.ID))
```
Take a look at the buffer and bismal point data up close (in Okinawa) using ggplot so as to label
```{r }
o.e <- extent(raster(xmn = -100, xmx = 600, ymn = 2000, ymx = 3500))

oki_buf<-st_crop(IR_buffer, o.e)
oki_bis<-st_crop(bis_km, o.e)

```

```{#r map plot oki buffers and bis samples}

plot(st_geometry(oki_buf), col = sf.colors(12, categorical = TRUE), border = 'grey', 
     axes = TRUE)
plot(st_geometry(st_centroid(IR_km)), pch = 3, col = 'red', add = TRUE)

plot(st_geometry(st_centroid(oki_bis)), cex = 0.2, pch = 16, col = 'orange', add = TRUE)


```

Can see that there is some clustering of Bismal samples inside vent sites as is to be expected from targeted sampling of known vent sites.

Take a closer look 

Plot wth ggplot and ggsflabel (https://yutannihilation.github.io/ggsflabel/)
```{r Map N. Oki. vent buffer and full Bis. samples}
#devtools::install_github("yutannihilation/ggsflabel")
library(ggsflabel)
library(gghighlight)

#Trim the bathymetry data to only include contours at selected depths eg. 150m
bf_150<- subset(bf_km, z  == seq(-100, -2000,- 150))

# make plot
ggplot()+



  # add vent buffers
  geom_sf(data = oki_buf) + 

    # add bismal samples 
  geom_sf(data = oki_bis, colour = "orange", 
             stroke = .5, size = 0.1, 
             alpha = 1, shape = 21) + 
  

 #add sample labels but be careful not to include too many samples
  geom_sf_label_repel(data = oki_buf, aes(label = Name.ID),
                        force = 100, nudge_x = -2) +

   #focus on a region of interest e.g. N.Okinawa site
    coord_sf(xlim = c(200,400), ylim = c(2900, 3100), expand = FALSE) +
  
    theme_classic() +
  
  labs(
    x = "Long.",              # x axis title
    y = "Lat."  # y axis title

  )+

 
  
  ggtitle(" N.Okinawa InterRidge Vents 5km buffer and Bismal samples")+
    theme(plot.title = element_text(hjust = 0.5))
```

Plot a particular vent site of interest, for example Yokosuka

```{r Map Yukosuka vent buffer and labelled Bis. samples}
#devtools::install_github("yutannihilation/ggsflabel")
library(ggsflabel)
library(gghighlight)

#Trim the extent of the data-sets otherwise there will be a lot of data and the plot will take some time

y.e <- extent(raster(xmn = -30, xmx = 50, ymn = 2790, ymx = 2810))
yuko_buf<-st_crop(IR_buffer, y.e)
yuko_bis<-st_crop(bis_km, y.e)


# make plot
ggplot()+

  
  # add vent buffers
  geom_sf(data = yuko_buf) + 

    # add bismal samples 
  geom_sf(data = yuko_bis, colour = "orange", 
             stroke = .5, size = 0.1, 
             alpha = 1, shape = 21) + 
  
  #add sample labels but be careful not to include too many samples
  geom_sf_label_repel(data = yuko_bis, aes(label = scientificName),
                        force = 100, nudge_x = -2, seed = 10, size = 2) +

  #add vent labels
  geom_sf_label(data = yuko_buf, aes(label = Name.ID), size = 3, fontface = "bold",nudge_y = 3 ,nudge_x = 8) +

   #focus on a region of interest e.g. Yokosuka site
    coord_sf(xlim = c(-30,48), ylim = c(2770, 2840), expand = FALSE) +
  
    theme_classic() +
  
  labs(
    x = "Long.",              # x axis title
    y = "Lat."  # y axis title

  )+

 
  
  ggtitle(" Yukosuka  Vents 5km buffer and Bismal samples Labelled")+
    theme(plot.title = element_text(hjust = 0.5))
```
I selected Yokusuka because I could see BISMaL samples within the buffer but they were not included in the final data-set. It turns out that the samples are described at too low a taxonomic resolution to be included. Below you will see that I only include samples identified to species level


Save this data-set of all bismal occurences within the vicinity of hydrothermal vents
```{r}
write.csv(IR_bis, "output/bis_interridge_5kmbuffer.csv")
```



# Trim Taxa by SFDVENT



Remove all occurences from BISMAL that aren't of vent species identified by sfdvent (Chapman *et al.,* 2019), described at the time as the most comprehensive database of vent-associated species
```{r merge by taxon}

datum<-merge(taxa, IR_bis, by.x = "Taxon",by.y =  "scientificName")
str(factor(datum$Taxon))
summary(unique(datum$Taxon))
```

Now do the same again but after **translating both BISMAL and SFDVENT species to WORMS standardised names** using the WORMS website (https://www.marinespecies.org/aphia.php?p=match) 

After running it through WoRMS the first time I found that some non-matched species were due to spelling mistakes (which WoRMS can't detect).

After finding the spelling mistakes, change them manually befor exporting them again
```{r }

#Planktotalea lamellibrachiae -> No Known Species Name

#Parasphingopyxis lamellibrachiae -> No Known Species Name

#Psychrobium conchae ->  No Known Species Name

#Truro coronatus coronatus -> Turbo coronatus
IR_bis$scientificName <- gsub('Truro coronatus coronatus', 'Turbo coronatus', IR_bis$scientificName)

#Oncaea conifera coronatus -> Oncaea furcula 
IR_bis$scientificName <- gsub('Oncaea conifera coronatus', 'Oncaea furcula ', IR_bis$scientificName)


```

Write the unique species names of the BISMaL and sFDvent database as .xlsx format as that is most easily read by the WoRMS Taxon match website. Also for some reason it works better on Safari than chrome...

```{r write files for translation in WoRMS}

# xlsx format is better with the worms matching software
library(openxlsx)

write.xlsx(unique(IR_bis$scientificName), "output/bis_IRbuffer_sp.xlsx", row.names = FALSE)
write.xlsx(unique(taxa$Taxon), "output/sfdvent_sp.xlsx", row.names = FALSE)

```

 
**After running the species lists through WoRMS** import the new datasets with 'accepted names' from where you saved them
```{r IN PROCESS}
w_IR_bis_sp<-read_excel("output/bis_irbuffer_sp_matched.xlsx")
w_taxa_sp <-read_excel("output/sfdvent_sp_matched.xlsx")

```
Combine accepted names with occurence records
```{r}
#Match all the accepted species names from the WoRMS translated BISMaL species list to the BISMaL occurences 
w_IR_bis<-merge(IR_bis, w_IR_bis_sp, by.x = "scientificName", by.y = "ScientificName...1")
```

Now all BIS occurences have a matching WorMS accepted name, match them to the SFdvent accepted names. And check how many unique vent sites and species we are left with.

```{r}
#Merge accepted bis names with accepted WoRMS names to effectively remove all BIS occurences that don't match
datum1<-merge( w_IR_bis, w_taxa_sp, by.x = "ScientificName_accepted",by.y =  "ScientificName_accepted")

#Translate species name to a factor from a character
str(factor(datum1$ScientificName_accepted))
#Check how many unique species we are left with
summary(unique(datum1$ScientificName_accepted))

#Translate vent name to a factor from a character
str(factor(datum1$Name.ID))
#Check how many unique vent sites we are left with
summary(unique(datum1$Name.ID))

```

Trim and save this dataset that represents all BISMaL occurences of species that match those identified as vent-associated in the SFDVENT database (once both are converted to accepted names through WoRMS) and are found within 5km of a vent site registered on the InterRidge database (ver.3.4)

From this point on the column 'species' will contain the WoRMS accepted name while the previous name used in the BISMaL data-set will be retained as 'bis_scientificName' for reference
```{r }

#Filter out additional columns created in the merging process
datum1 <- datum1[c ("ScientificName_accepted","family", "genus","specificEpithet","scientificName", "locality", "geometry", "eventDate", "fieldNumber", "institutionCode", "datasetName", "basisOfRecord", "references", "occurrenceID", "catalogNumber", "recordNumber", "recordedBy", "BISMaLtaxonID", "Name.ID")]

#Re-name the columns 
names(datum1)<-list("species","bis_family", "bis_genus","bis_specificEpithet","bis_scientificName", "locality", "geometry", "eventDate", "fieldNumber", "institutionCode", "datasetName", "basisOfRecord", "references", "occurrenceID", "catalogNumber", "recordNumber", "recordedBy", "BISMaLtaxonID", "vent_IR")

#Re-integrate the interridge data from the original data-set
bis_ir_sfd_vent<-merge(datum1, vent, by.x ="vent_IR", by.y = "Name.ID" )

write.csv(bis_ir_sfd_vent,"output/bis_by_IR_sfdvent.csv")
```

# Add occurences from Literature

Import species by site matrices from various sources, transform them into lists of occurences, combine them with the above 'BISMaL-by-InterRidge-by-SFdvent' data-set and then transform them into a site-by-species matrix

.... 

## Tunnicliffe (*unpublished*)

Sent on 2020/03/20 from Verena Tunnicliffe: "All but one record derives from our collections on NOAA sponsored Ring of Fire expeditions (2004, 2006, 2014). Where no 'confirmed' but a species ID - was by me by morphology; most are not difficult to identify.  If unsure, I use 'cf'. update from Watanabe et al 2019"

```{r}
#Import
tun_origi<-read_excel("data/Tunnicliffe_Arc Data for Otis Brunner OIST_update.xlsx")

#Remove columns and rows outside of the matrix
tun<-tun_origi[,4:13]
tun<-tun[5:44,]
colnames(tun)<-tun[1,]
tun<-tun[-1,]

#'Melt' the matrix so that every species is aligned to every vent with a third column that specifies if it is present (1) or absent(0)
tun.o<-reshape2:::melt(tun, id = "Species")

#Remove all absences
tun.o<- tun.o %>% 
  filter(value == "1") 

```

Now combine this species at vent list with the bismal dataset
```{r}
#rearrange column order to match bis data
tun.o1<-tun.o[c("variable","Species")]
#bind the rows of tun and bis by adding columns first
tun.o1[,3:length(bis_ir_sfd_vent)] <- c(NA)
names(tun.o1)<-names(bis_ir_sfd_vent)

#Add the reference of this dataset to its rows
tun.o1$references<- "Verena Tunnicliffe, unpublished"

#Add to the existing dataset
datum<-rbind(as.data.frame(bis_ir_sfd_vent),tun.o1)
```



## Giguere (*unpublished*)

Sent on 2020/05/30 from Thomas Giguere

```{r}
#Import
gig_origi<-read_excel("data/Giguere_Mariana BASC Vent PA List.xlsx")

#Remove rows and rename columns
gig<-gig_origi[1:38,]
gig_origi[39,1]<-"Species"
colnames(gig)<-gig_origi[39,]


#'Melt' the matrix so that every species is aligned to every vent with a third column that specifies if it is present (1) or absent(0)
gig.o<-reshape2:::melt(gig, id = "Species")

#Remove all absences
gig.o<- gig.o %>% 
  filter(value == "1") 

```

Now combine this species at vent list with the bismal dataset
```{r}
#rearrange column order to match bis data
gig.o<-gig.o[c("variable","Species")]
#bind the rows of tun and bis by adding columns first
gig.o[,3:length(bis_ir_sfd_vent)] <- c(NA)
names(gig.o)<-names(bis_ir_sfd_vent)

#Add the reference of this dataset to its rows
gig.o$references<- "Thomas Giguere, unpublished"

#Add to the existing dataset
datum<-rbind(as.data.frame(datum),gig.o)

```


##Nakajima *et al*., 2014

Incorporate Table S1 from Nakajima *et al*., 2014
"Table S1. Benthic animal datasets for 155 species from 42 deep-sea chemosynthetic ecosystems around the Japanese Archipelago"

```{r}
#Import
naka_origi<-read_excel("data/Nakajima_ddi12204-sup-0001-tables1.xlsx")

#Remove rows and rename columns
naka<-naka_origi[,5:52]
naka<-naka[4:163,]
naka_origi[3,5]<-"Species"
colnames(naka)<-naka_origi[3, 5:52]

#Remove non-vent sites
naka<-cbind(naka[,1],naka[,33:48])
  
#Remove NA column
naka<-naka[,-11]

#Transform + into 1
naka[naka == "+"] <-"1"

#'Melt' the matrix so that every species is aligned to every vent with a third column that specifies if it is present (1) or absent(0)
naka<-reshape2:::melt(naka, id = "Species")

#Remove all absences
naka<- naka %>% 
  filter(value == "1") 

```

Now combine this species at vent list with the bismal dataset
```{r}
#rearrange column order to match bis data
naka<-naka[c("variable","Species")]
#bind the rows of naka and bis by adding columns first
naka[,3:length(bis_ir_sfd_vent)] <- c(NA)
names(naka)<-names(bis_ir_sfd_vent)

#Add the reference of this dataset to its rows
naka$references<- "Nakajima et al., 2014"

#Add to the existing dataset
datum<-rbind(as.data.frame(datum),naka)
```


##Watanabe *et al*., 2019

Data was ripped from Table S2 from Watanabe et al., 2019.

```{r}
#Import
watanabe_origi<-read_excel("data/Watanabe.2019_Table S2.xlsx")

#Rename species to replace '.' seperator with blank space
watanabe_origi$Species <- gsub(".", " ",fixed=TRUE, watanabe_origi$Species)


#'Melt' the matrix so that every species is aligned to every vent with a third column that specifies if it is present (1) or absent(0)
wata<-reshape2:::melt(watanabe_origi, id = "Species")

#Remove all absences
wata<- wata %>% 
  filter(value == "1") 

```

Now combine this species at vent list with the bismal dataset
```{r}
#rearrange column order to match bis data
wata<-wata[c("variable","Species")]
#bind the rows of wata and bis by adding columns first
wata[,3:length(bis_ir_sfd_vent)] <- c(NA)
names(wata)<-names(bis_ir_sfd_vent)

#Add the reference of this dataset to its rows
wata$references<- "Watanabe et al., 2019"

#Add to the existing dataset
datum<-rbind(as.data.frame(datum),wata)
```

Write complete occurence list
```{r}

write.csv(datum,"output/occurence_BisNakTunWatGig.csv")

#Write unique species to check against WORMS
write.csv(unique(datum$species),"output/species_BisNakTunWatGig.csv")
```


# Reformat Occurence List into Site-by-Species Matrix

There are multiple individual species that are not confirmed at the species level, these can be targetted and removed from the dataset by targetting key abbreviations that are used instead of species names.
```{r}

#Remove occurences that are not to species level
occur<-datum[!grepl("sp", datum$species),]
occur<-occur[!grepl("aff", occur$species),]
occur<-occur[!grepl("cf", occur$species),]
occur<-occur[!grepl("different", occur$species),]
occur<-occur[grepl(" ", occur$species),]
occur<-occur[grepl(".", occur$species),]

#List all species names to check for abnormalities
unique(occur$species)
```
Check for species name spelling mistakes as WoRMS cannot detect them. Running the species list through worms will help detect these spelling mistakes but then go back and change them before running the species list through WoRMS once again

After finding the spelling mistakes, change them manually (this includes three-part names which it also can't recognize)
```{r}


#Alvinicaris marimonte -> Alvinocaris marimonte 
occur$species <- gsub('Alvinicaris marimonte', 'Alvinocaris marimonte', occur$species)

# desbruyeresi nipponensis <- Eosipho nipponensis
occur$species <- gsub('desbruyeresi nipponensis', 'Eosipho nipponensis', occur$species)

#Neptunea insularis robusta <- Neptunea insularis
occur$species <- gsub('Neptunea insularis robusta', 'Neptunea insularis', occur$species)


#Use this script to detect if the names are in the dataset
#subset(occur, species == "")
```

Export and check this species list agains WoRMS 
```{r}

write.xlsx(unique(occur$species), "output/BisNakWatTunGig_sp.xlsx", row.names = FALSE)
```


There are some changes to species names according to WoRMS so time to translate them again

```{r}

w_occur<-read_excel("output/bisnakwattungig_sp_matched.xlsx")
```

Replace the names with WoRMS accepted names
```{r}
#Match all the accepted species names from the WoRMS translated combined species list to the combined occurences 
w_occur_acc<-merge(occur, w_occur, by.x = "species", by.y = "ScientificName...1")

```

Transform occurence at site dataframe into a site-by-species matrix.
```{r}
#Transform into matrix
mat<-splist2presabs(w_occur_acc, "vent_IR", "species", keep.n = FALSE)

#Check for all unique Vent site names
unique(mat$vent_IR)

#make sure to include vent site name as the row name
rownames(mat)<-mat[,1]
mat<-mat[,-1]
mat<-as.matrix(mat)
```
**Many of the vent site names are alliases of one another.** It seems we have 58 vent sites but we have less because the alias rows need to be merged. For standardisation, the vent name recognised by InterRidge will be kept and the aliases used in the other data-sets will be removed


```{r}
#Combine Bayonaise
mat["Bayonnaise Knoll caldera" , ] <- mat["Bayonnaise Knoll caldera", ] + mat["Bayonaise", ]
mat<-mat[rownames(mat) != "Bayonaise", ]

	
#Combine Daikoku
mat["Daikoku volcano", ] <- mat["Daikoku volcano", ] + mat["Daikoku", ]
mat<-mat[rownames(mat) != "Daikoku", ]

#Combine 'Iheya' from Nakajima et al., 2014 with the InterRidge accepted name  'Sakai'
mat["Sakai", ] <- mat["Sakai", ] + mat["Iheya", ]
mat<-mat[rownames(mat) != "Iheya", ]

#Combine Iheya North
mat["North Knoll, Iheya Ridge", ] <- mat["North Knoll, Iheya Ridge", ] + mat["Iheya North", ]
mat<-mat[rownames(mat) != "Iheya North", ]

#Combine Alice Springs
mat["Alice Springs Field", ] <- mat["Alice Springs Field", ] + mat["Illium-Alice Springs", ]
mat<-mat[rownames(mat) != "Illium-Alice Springs", ]

#Combine 'Izena' from Nakajima et al., 2014 with the InterRidge accepted name  'Izena Cauldron'
mat["Izena Cauldron", ] <- mat["Izena Cauldron", ] + mat["Izena", ]
mat<-mat[rownames(mat) != "Izena", ]

#Combine 'Kaikata' from Watanabe et al., 2019 with the InterRidge accepted name  'Kaikata Seamount'
mat["Kaikata Seamount", ] <- mat["Kaikata Seamount", ] + mat["Kaikata", ]
mat<-mat[rownames(mat) != "Kaikata", ]

#Combine 'Kasuga 2' from Tunnicliffe et al., unpublished with the InterRidge accepted name  'Kasuga 2 Seamount'
mat["Kasuga 2 Seamount", ] <- mat["Kasuga 2 Seamount", ] + mat["Kasuga 2", ]
mat<-mat[rownames(mat) != "Kasuga 2", ]

#Combine 'Kasuga II' from Nakajima et al., 2014 with the InterRidge accepted name  'Kasuga 2 Seamount'
mat["Kasuga 2 Seamount", ] <- mat["Kasuga 2 Seamount", ] + mat["Kasuga II", ]
mat<-mat[rownames(mat) != "Kasuga II", ]

#Combine 'Kasuga2' from Tunnicliffe et al., unpublished with the InterRidge accepted name  'Kasuga 2 Seamount'
mat["Kasuga 2 Seamount", ] <- mat["Kasuga 2 Seamount", ] + mat["Kasuga2", ]
mat<-mat[rownames(mat) != "Kasuga2", ]

#Combine 'Minami-Ensei' from Nakajima et al., 2014 with the InterRidge accepted name  'Minami-Ensei Knoll'
mat["Minami-Ensei Knoll", ] <- mat["Minami-Ensei Knoll", ] + mat["Minami-Ensei", ]
mat<-mat[rownames(mat) != "Minami-Ensei", ]

#Combine 'Mokuyo' from Watanabe et al., 2019 with the InterRidge accepted name  'Mokuyo Seamount'
mat["Mokuyo Seamount", ] <- mat["Mokuyo Seamount", ] + mat["Mokuyo", ]
mat<-mat[rownames(mat) != "Mokuyo", ]

#Combine 'Myojin' from Watanabe et al., 2019 with the InterRidge accepted name  'Myojin Knoll'
mat["Myojin Knoll", ] <- mat["Myojin Knoll", ] + mat["Myojin", ]
mat<-mat[rownames(mat) != "Myojin", ]

#Combine 'Myojin Sho' from Watanabe et al., 2019 with the InterRidge accepted name  'Myojinsho Caldera'
mat["Myojinsho Caldera", ] <- mat["Myojinsho Caldera", ] + mat["Myojin Sho", ]
mat<-mat[rownames(mat) != "Myojin Sho", ]

#Combine 'Nikko' from Watanabe et al., 2019 with the InterRidge accepted name  'Nikko volcano'
mat["Nikko volcano", ] <- mat["Nikko volcano", ] + mat["Nikko", ]
mat<-mat[rownames(mat) != "Nikko", ]

#Combine 'Eifuku' from Nakajima et al., 2014 with the InterRidge accepted name  'Northwest Eifuku'
mat["Northwest Eifuku", ] <- mat["Northwest Eifuku", ] + mat["Eifuku", ]
mat<-mat[rownames(mat) != "Eifuku", ]

#Combine 'NW Eifuku' from _____ with the InterRidge accepted name  'Northwest Eifuku'
mat["Northwest Eifuku", ] <- mat["Northwest Eifuku", ] + mat["NW Eifuku", ]
mat<-mat[rownames(mat) != "NW Eifuku", ]

#Combine 'NW Rota' from Tunnicliffe et al., unpublished with the InterRidge accepted name  'Northwest Rota-1 volcano'
mat["Northwest Rota-1 volcano", ] <- mat["Northwest Rota-1 volcano", ] + mat["NW Rota", ]
mat<-mat[rownames(mat) != "NW Rota", ]

#Combine 'Urashima-Pika' from Giguere et al., unpublished with the InterRidge accepted name  'Pika'
mat["Pika", ] <- mat["Pika", ] + mat["Urashima-Pika", ]
mat<-mat[rownames(mat) != "Urashima-Pika", ]

#Combine 'Hatoma' from Nakajima et al., 2014 with the InterRidge accepted name  'SPOT, Hatoma Knoll'
mat["SPOT, Hatoma Knoll", ] <- mat["SPOT, Hatoma Knoll", ] + mat["Hatoma", ]
mat<-mat[rownames(mat) != "Hatoma", ]

#Combine 'Yonaguni IV' from Nakajima et al., 2014 with the InterRidge accepted name  'SPOT, Yonaguni Knoll IV'
mat["SPOT, Yonaguni Knoll IV", ] <- mat["SPOT, Yonaguni Knoll IV", ] + mat["Yonaguni IV", ]
mat<-mat[rownames(mat) != "Yonaguni IV", ]

#Combine 'Suiyo' from Watanabe et al., 2019 with the InterRidge accepted name  'Suiyo Seamount'
mat["Suiyo Seamount", ] <- mat["Suiyo Seamount", ] + mat["Suiyo", ]
mat<-mat[rownames(mat) != "Suiyo", ]

#Combine 'Sumisu' from Nakajima et al., 2014 with the InterRidge accepted name  'Sumisu Caldera'
mat["Sumisu Caldera", ] <- mat["Sumisu Caldera", ] + mat["Sumisu", ]
mat<-mat[rownames(mat) != "Sumisu", ]

#Check the unique vent names we are left with
unique(rownames(mat))
```

Adjust the name of vent sites to match names recognised in InterRidge
```{r}

#Rename Chamorro as Chamorro volcano
rownames(mat)[rownames(mat) == "Chamorro"] <- "Chamorro volcano"

#Rename E Diamante as East Diamante volcano
rownames(mat)[rownames(mat) == "E Diamante"] <- "East Diamante volcano"

#Rename SeamountX as Seamount X
rownames(mat)[rownames(mat) == "SeamountX"] <- "Seamount X"

```

Take a look at the number of unique species (sp. richness) we have at each vent

```{r}

#Ensure this data-set is only presence/absence. Transform all values <1 (resulting from merging allias vent sites) into 1
mat[mat>1]<-1

#Take a llok at the species richness (row sum) of each vent site
cbind(rownames(mat),rowSums( mat ))
       
```

Now add back in the metadata from InterRidge

First add in metadata for vents not recognized as sites in InterRidge
```{r}

#Site from Giguere et al. (unpublished) 
#metadata from https://www.utdallas.edu/gmtl/research/submarine-vents-in-mariana-forearc-arc-backarc-system/archaean-vent/
Archaean<-data.frame("Archaean", 12.939, 143.632, "Mariana Trough", 3060, "back-arc spreading center")
names(Archaean)<-names(vent)

#Site from Giguere et al. (unpublished) 
#metadata from http://www.godac.jamstec.go.jp/catalog/data/doc_catalog/media/shinkai13_06.pdf
Burke<-data.frame("Burke", 18.109, 144.432, "Mariana Trough", 3660, "back-arc spreading center")
names(Burke)<-names(vent)

vent<-rbind(vent, Archaean, Burke)
```

Now that the vent site names match thesite-by-species data can be combined with the site metadata

```{r}

#Bring the rownames back as a column
mat<-cbind(rownames(mat), mat)

#Add the InterRidge Metadata
mat<-merge(mat, vent, by.x = "V1", by.y = "Name.ID")

#Rename the vent site column
colnames(mat)[1] <- "vent"
```


Some Final Manual Adjustments:

Forecast is mistankely catagorized as an arc vent but is in fact a back-arc vent (Tunnicliffe, personal communication)

```{r}

#Give the Forecat Region setting the same as Alice Springs
mat[8,114]<-mat[1,114]

#Give the Forecat Region the same as Alice Springs
mat[8,116]<-mat[1,116]

```

**And we're done!** Save the site-by-species matrix 
```{r}
#Save to the output sub-folder of this r-project
write.csv(mat, "output/site_by_species.csv", row.names = FALSE)
```
